---
title: "p8105_hw1_mp3653"
author: "Matthew Perrotta"
date: "September 18, 2018"
output: github_document
---
Load Tidyverse package
```{r tidyverse}
library(tidyverse)
```

_Problem 1_

Create a data frame

```{r create dataframe}
set.seed(1)

df = tibble(
  vec_numeric = runif(10, 1, 5),
  vec_logical = vec_numeric > 2,
  vec_char = c('Malaria', 'Ebola Virus', 'Influenza', 'MRSA', 'Poliovirus', 'Dengue', 'E. Coli', 'Measles', 'VZV', 'MERS'),
  vec_factor = factor(c('Parasite', 'Virus', 'Virus', 'Bacteria', 'Virus', 'Virus', 'Bacteria', 'Virus', 'Virus', 'Virus'))
)

df
```

Take the mean of each variable
```{r means}
mean(df$vec_numeric)
mean(df$vec_logical)
mean(df$vec_char)
mean(df$vec_factor)
```
the mean function provides a numeric output for the numeric and logical vectors and provides a NA output for the character and factor vectors. The numeric and logical vectors contain values that can be used in a mathematical function (the logical vector's TRUE/FALSE values can be seen as 1's and 0's) while the character and factor vectors contain non-numeric values.

Coercing logical, character, and factor variables to numeric
```{r eval=FALSE}
as.numeric(df$vec_logical)
as.numeric(df$vec_char)
as.numeric(df$vec_factor)
```
 >When coercing the logical, character, and factor vectors the following happened:
 
  * The logical vector became a series of 1's and 0's, with 1's corresponding to true statements and 0's  corresponding to false statements.
  * The character vector became a series of NA's
  * The factor vector became numeric values 1 through 3, with each number assigned to a specific value (1 = Bacteria, 2 = Parasite, 3 = Virus; number assignment seems to be alphabetical)

```{r eval=FALSE}
as.numeric(as.factor(df$vec_char))
as.numeric(as.character(df$vec_factor))
```
When the character vector was coerced to a factor vector, the values were assigned numeric levels by alphabetical order. When this factor vector was coerced to a numeric vector these numeric levels were outputted in the original order the character values were in.

When the factor vector was coerced to a character vector, the each value became a string of characters. When this character vector was coerced to a numeric vector the output was a series of NA's, as character strings cannot be coerced to numbers.

***
_Problem 2_


